---
title: "Dilution Gauging"
author: "Karin Emanuelson"
date: "11/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(knitr)
library(kableExtra)
library(here)
```

### Read in field data and enter metadata
```{r}
# Read in data from datalogger
cond <- read.csv("data/corrected conductivity.csv", skip = 3, header = T)[ ,1:6]

names(cond) <- c("Timestamp", "time_s", "RN", "Cond_cor_mS", "Ct_raw_mS", "Temp_C")

#clean
cond<- cond %>% 
  mutate(time_s= as.numeric(time_s)) %>% 
  mutate(Timestamp = as.POSIXct(Timestamp,format="%H:%M:%S"))

# Read in metadata
## reach distance (same for all trials)
dist <- 20
## injection 1
mass1<-223.8
injTime1 <- as.POSIXct("15:11:35", format = "%H:%M:%S")
## injection 2
mass2<-418.4
injTime2 <- as.POSIXct("15:21:40", format = "%H:%M:%S")
## injection 3
mass3<-312.6
injTime3 <- as.POSIXct("15:41:45", format = "%H:%M:%S")
```

### Quick plot to look at data
```{r}
ggplotly(ggplot(data = cond) +
  geom_point(aes(x = time_s, y = Cond_cor_mS)))
```

### Determine begin, end, and rise time for each slug injection.
#### Use the plotly above to hover mouse and determine the begin time (earliest time to use for background correction), rise time (latest time to use for background correction jsut before the conductivity increases), and end time (time at which the conductivity is back to background concentrations).
```{r}
# variables for slug 1
rise <- as.numeric(2105)
begin <- as.numeric(1910)
end <- as.numeric(2410)

# variables for slug 2
rise2 <- as.numeric(3115)
begin2 <- as.numeric(2875)
end2 <- as.numeric(3675)

# variables for slug 3
rise3 <- as.numeric(3785)
begin3 <- as.numeric(3690)
end3 <- as.numeric(4340)
```

### Separate each slug breakthrough curve (BTC) for analysis 
```{r}
#slugs
slug1 <- filter(cond, time_s>begin& time_s<end) %>% 
    mutate(time_since_injection = as.numeric(difftime(Timestamp, injTime1, units='secs'))) %>% # Create time since injection column
    filter(time_since_injection >=0) # filter to only include time since injection
slug2 <- filter(cond, time_s>begin2& time_s<end2) %>% 
    mutate(time_since_injection = as.numeric(difftime(Timestamp, injTime2, units='secs'))) %>% # Create time since injection column
    filter(time_since_injection >=0) # filter to only include time since injection
slug3 <- filter(cond, time_s>begin3& time_s<end3) %>% 
    mutate(time_since_injection = as.numeric(difftime(Timestamp, injTime3, units='secs'))) %>% # Create time since injection column
    filter(time_since_injection >=0) # filter to only include time since injection
```

### Complete dilution gauging analysis for each slug
Determine baseline for slug 1, 
```{r}
# baseline for slug 1
start.inj.idx<- which(slug1$time_s==rise)
baseline <- median(slug1$Cond_cor_mS[0:(start.inj.idx)])

# set dt (the timestep of data collection)
dt <- slug1$time_s[2] - slug1$time_s[1]

# background correct
slug1 <- slug1 %>%
  mutate(bgcor_con = Cond_cor_mS - baseline) %>% # background correct Conductivity
  mutate(bgcor_con =ifelse(bgcor_con < 0, 0, bgcor_con)) %>% # ensure background correction is not less than zero
  mutate(NaCl = bgcor_con/2) %>% # convert from conductivity (mS/cm) to concentration NaCl (g/L)
  mutate(IntC_s1 = NaCl+lag(NaCl,default=0))%>% # first step of integration 
  mutate(IntC = cumsum(IntC_s1*.5*dt))%>% # complete integration
  mutate(recovery = IntC/max(IntC)) # Calculate recovery

# Determine flow
Q1<-mass1/max(slug1$IntC)

# Determine modal, 5%, 50% and 95% velocities
## Modal velocity
peak1.idx<- which.max(abs(slug1$bgcor_con)) # find the index of the point that has the max concentration
peak1_time <- as.numeric(difftime(slug1$Timestamp[peak1.idx], injTime1, units='secs')) # Calculate the time from injection to that max point in seconds
modalvel1 <- dist/peak1_time # determine velocity by dist/time

## 5 percent recovery velocity
fiveRecovery.idx1<- which.min(abs(slug1$recovery-.05)) # find the index of the point that has recovery equal to .05 (or 5%)
fiveper_time1 <- as.numeric(difftime(slug1$Timestamp[fiveRecovery.idx1], injTime1, units='secs')) # Calculate the time from injection to that 5% recovery point in seconds
Fivevel1 <- dist/fiveper_time1 # determine velocity by dist/time

## 50% recovery velocity
halfRecovery.idx1<- which.min(abs(slug1$recovery-.50)) # find the index of the point that has recovery equal to .5 (or 50%)
fiftyper_time1 <- as.numeric(difftime(slug1$Timestamp[halfRecovery.idx1], injTime1, units='secs')) # Calculate the time from injection to that 50% recovery point in seconds
halfvel1 <- dist/fiftyper_time1 # determine velocity by dist/time

## ninety-five revovery velocity
ninetyfiveRecovery.idx1 <- which.min(abs(slug1$recovery-.95)) # find the index of the point that has recovery equal to .95 (or 95%)
ninetyfiveper_time1 <- as.numeric(difftime(slug1$Timestamp[ninetyfiveRecovery.idx1], injTime1, units='secs')) #Calculate the time from injection to that 95% recovery point in seconds
ninetyfivevel1 <- dist/ninetyfiveper_time1 # determine velocity by dist/time


# Create table of Q and velocities
slug1_table<- data.frame(Q1, modalvel1, Fivevel1, halfvel1, ninetyfivevel1)
names(slug1_table) <- c("Q (L/s)", "Modal Velocity (m/s)", "5% Velocity (m/s)", "50% Velocity (m/s)", "95% Velocity (m/s)")
print(slug1_table)
```

### Graph BTC and Recovery
```{r}

# slug1 plot
ggplot(data = slug1)+
  geom_point(aes( x = time_since_injection, y = NaCl))+
  labs(x = "Time since Injection (sec)", y = "Background Corrected NaCl (g/L))")+
  theme_bw()

# normalized slug1 plot
ggplot(data = slug1)+
  geom_point(aes( x = time_since_injection, y = recovery))+
  labs(x = "Time since Injection (sec)", y = "% of NaCl")+
  theme_bw()

```

