---
title: "Velocity area discharge"
author: "tim covino"
date: "9/7/2021"
output: html_document
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, results = FALSE)

```

```{r}
library(tidyverse)
library(plotly)
```
### This part you will need to change. Remember that it is best practice to work from projects. If you have a WR696 project with sub-folders you would open the project then navige to the velo_area (for example) folder and open your R markdown. The reason this is nice is that project will open with WR696 as the "working directory". Problems loading data are generally due to thinking you are in a different wd than you actually are. Projects essentially eliminate this issue. 
```{r}
cross <- read_csv("data/velo_area_data_R.csv")
```

```{r}

cross <- cross %>% 
  mutate(depth_m = depth_cm/100)


cross <- cross %>% 
  mutate(area_m2 = (lead(cross$width_m) - lag(cross$width_m)) / 2 *
  cross$depth_m, q = area_m2 * velo_m_s)

tot_q <- sum(cross$q, na.rm = TRUE)

```

```{r}
cross_percent <- cross %>% 
  mutate(q_per = (q/tot_q)*100)

sections_exceeding <- cross_percent %>% 
  filter(q_per > 5)

mean_velo <- cross %>% 
  summarize(mean_velo = mean(velo_m_s))

total_area <- cross %>% 
  summarize(total_area = sum(area_m2, na.rm = TRUE))
```

```{r}
fig <- cross %>% 
  mutate(neg_depth = depth_m * -1) %>% 
  ggplot(aes(x = width_m, y = neg_depth)) +
  geom_point() +
  geom_line() +
  ylab("Depth (m)") +
  xlab("Width (m)") +
  theme_linedraw(base_size = 16) +
  ylim(-15, 0)

ggplotly(fig)

```


